---
title: "Factors affecting multiple paternity: insights from the eastern quoll"
author: "Brockett B M"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  html_document:
    toc: true
    number_sections: true
    top_depth: 3
    toc_float:
      collapsed: true
    theme: cerulean
    highlight: pygments
    self_contained: true
editor_options:
  chunk_output_type: console
knit: "(function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), 'tutorial.html')) })"
---

# **Background**

This [R project](https://support.posit.co/hc/en-us/articles/200526207-Using-RStudio-Projects) presents the datasets and workflows used in the *Factors affecting multiple paternity: insights from the eastern quoll (*Dasyurus viverrinus*)*  project. For more information, see the [GitHub repo](https://github.com/coexistence-conservation-lab/mfws-eq-breeding-system) README.

The purpose of this modelling is to investigate what is affecting breeding success of the eastern quoll (*Dasyurus viverrinus*), in an effort to understand drivers of reproductive sucess and population-level patterns of genetic diversity. Prior investigation using single-nucleotide polymorphisms (genetic data) has provided us with a pedigree for all known individuals at [Mulligans Flat Woodland Sanctuary](https://www.mulligansflat.org.au/), from which we have been able to determine some reproductive patterns. What remains uncertain is: what makes a 'good' eastern quoll (i.e., what traits lead to higher reproductive success/output)?

We posed the question: *"What affects breeding success in  population"* using a hurdle-model approach.
  
  1. Did **morphological** (e.g., pelage colour morph) or **genetic** (e.g., heterozygosity) factors drive breeding success? 
  2. For individuals that breed successfully, do these factors influence the number of offspring produced?

Note that in this analysis "successful" reproduction is equivalent to recruiting offspring into the next generation, rather than conception/birth being the endpoint. This is because individuals can only be captured and  microchipped/sampled if they survive from birth to dispersal from the natal den (and, therefore, they are captured in population monitoring).

# **Setup**

## Global options

When you [knit this document 🧶](https://rmarkdown.rstudio.com/authoring_quick_tour.html) (e.g., render to HTML), you can add **chunk options** to control what is shown in the output (e.g., `{r, message = FALSE, results = 'hide', warning = FALSE}`).

To avoid long chunk headers like the one above, use these custom options:  

  - **asis:** renders output *as-is* (e.g., for dynamic summaries)
  - **echo_only:** shows *code only* (hides results and messages)
  - **no_echo:** shows *output and messages* (hides code) 
  - **results_only:** shows *results only* (hides code and messages)

```{r global-options, echo = FALSE}
# Set default knitting options 
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)

# Set custom chunk knitting behaviours 
knitr::opts_hooks$set(
  
  # Render output as-is for chunks with "asis = TRUE"
  asis = function(x) {
   if (isTRUE(x$asis)) x$results <- "asis" 
   x}, 
  
  # Disable messages, results, and warnings for chunks with "echo_only = TRUE"
  echo_only = function(x) {
    if (isTRUE(x$echo_only)) {
      x$message <- FALSE
      x$results <- "hide"
      x$warning <- FALSE} 
    x},
  
  # Disable echo for chunks with "no_echo = TRUE"
  no_echo = function(x) {
    if (isTRUE(x$no_echo)) x$echo <- FALSE 
    x}, 
  
  # Disable echo, message, and warning for chunks with "results_only = TRUE"
  results_only = function(x) {
    if (isTRUE(x$results_only)) {
      x$echo    <- FALSE
      x$message <- FALSE
      x$warning <- FALSE} 
    x}) 
```

## Packages

  1. Dependancies 

The `ccl` package provides lab-specific functions used throughout this template, and `pacman` lets you install and load multiple packages at once. 

```{r, echo_only = TRUE}
# Install or update ccl package from GitHub 
remotes::install_github(
  "coexistence-conservation-lab/ccl", dependencies = FALSE)

# Load pacman (manages package installation and loading in one step)
library(pacman)

# Install BiocManager
if (!requireNamespace("BiocManager", quietly = TRUE))
  install.packages("BiocManager")

# Install SNPRelate and qvalue
BiocManager::install("SNPRelate")
BiocManager::install("qvalue")

# Install all missing and required packages for dartR (version 2.9.7)
gl.install.vanilla.dartR()
```

  2. Project-specific packages 

```{r, echo_only = TRUE}
# Install and load packages
pacman::p_load(
  
  # Data import and export 
  adegenet, dartR, data.table, hierfstat, 
  SNPRelate, openxlsx, readxl, writexl, yaml, 
  
  # Data wrangling 
  car, grid, gridExtra, gtable, janitor, tidyverse, 
  
  # Plotting 
  corrplot, ggeffects, ggpubr, patchwork, plotly, RColorBrewer, 
  scales, sjPlot, viridis, viridisLite, webshot, 
  
  # Reporting 
  english, kableExtra, rlang, rprojroot, 
  
  # Statistical modelling 
  AICcmodavg, bbmle, DHARMa, effects, glmmTMB, 
  lme4, lmtest, MuMIn, performance, trend, 
  
  # Utilities 
  beepr, cli, crayon, glue, here) 
```

```{r}
# Load ccl after all other packages to avoid masking conflicts
library(ccl)

# Prevent other packages (e.g., cowplot) from masking dplyr functions
select <- dplyr::select
recode <- dplyr::recode
```

## Resources

### Display abbreviations

```{r, eval = FALSE}
# Read in abbrevations list and print it as a clean table
(def <- print_clean_table(raw_data, sheet = "definitions"))
```

### Set constants

```{r}
# Define constants (e.g., repeated ggsave() arguments)  
dpi         <- 300
map_height  <- 6
map_width   <- 6
plot_height <- 6 
plot_width  <- 9

# Define MFWS centre coordinates 
mfws_lat    <- -35.166447
mfws_lon    <- 149.166040
```

### Load functions

```{r}
# Load project-specific functions from functions.R and report loaded objects
source(here::here("functions.R"))

# Check functions loaded by testing convert_date()
convert_date(c("2024-01-15", "45000", "2023-06-01 extra", "99999"))
```

# **Data preparation**

## SNP quality

```{r snp-setup}
# Set working directory
dartdata     <- "SNP_2_bettongANDduplicates_removed.csv"
ind_metadata <- "idpop_quoll_allinfo.csv"
gl.data      <- gl.read.dart(dartdata, ind.metafile = ind_metadata)

# Check data object column names
names(gl.data$other$loc.metrics)
```

```{r snp-quality-metrics}
# Mean depth across both alleles (default DArTSeq provides depth per allele)
depth_bothalleles <- gl.data$other$loc.metrics$AvgCountRef +
  gl.data$other$loc.metrics$AvgCountSnp

# Ratio of ref to SNP allele seq depth; used to check for PCR bias
# towards one allele, which may indicate null allele propensity
alleleDepthRatio <- gl.data$other$loc.metrics$AvgCountRef /
  gl.data$other$loc.metrics$AvgCountSnp
```

```{r snp-quality-plots}
glPlot(gl.data)
par(mfrow = c(2, 2))

hist(
  gl.data$other$loc.metrics$CallRate,
  xlab = "Call Rate", 
  main = "Call Rate")

hist(
  depth_bothalleles,
  xlab = "Depth", 
  main = "Depth")

hist(
  alleleDepthRatio,
  xlab = "Allele Seq Depth Ratio", 
  main = "Allele Seq Depth Ratio")

hist(
  gl.data$other$loc.metrics$RepAvg,
  xlab = "Repeatability", 
  main = "Repeatability")
```

## Population genetics

### Calculate metrics

```{r popgen-metrics}
# Overall allele frequencies, assuming bi-allelic SNPs
gl.data$other$loc.metrics$alf <- gl.alf(gl.data)[, 2]

# Allele frequencies and heterozygosities within sub-pops
allelefreq1 <- aggregate(
  gl.data, list(gl.data$pop),
  function(x) {(mean(x, na.rm = TRUE) / 2)})

hetcount <- aggregate(
  gl.data, list(gl.data$pop),
  function(x) {sum(x[x == "1"], na.rm = TRUE)})

samplsize <- aggregate(
  gl.data, list(gl.data$pop),
  function(x) {length(x[!is.na(x)])})

Hs  <- 1 - (allelefreq1[, -1]^2) - ((1 - allelefreq1[, -1])^2)
Ho  <- hetcount[, -1] / samplsize[, -1]
Fis <- (Hs - Ho) / Hs
```

```{r}
# Append means across sub-pops to genlight object
gl.data$other$loc.metrics$Ht      <- 2 * gl.data$other$loc.metrics$alf *
                                    (1 - gl.data$other$loc.metrics$alf)

gl.data$other$loc.metrics$meanHs  <- colMeans(Hs,              na.rm = TRUE)

gl.data$other$loc.metrics$meanHo  <- colMeans(Ho,              na.rm = TRUE)

gl.data$other$loc.metrics$meanFis <- colMeans(Fis,             na.rm = TRUE)

gl.data$other$loc.metrics$mean_p  <- colMeans(allelefreq1[-1], na.rm = TRUE)

gl.data$other$loc.metrics$Ho      <- colSums(hetcount[ , -1],  na.rm = TRUE) /
                                     colSums(samplsize[, -1])

gl.data$other$loc.metrics$Fit     <- (gl.data$other$loc.metrics$Ht - 
                                      gl.data$other$loc.metrics$Ho) / 
                                      gl.data$other$loc.metrics$Ht

# Estimate FST as (Ht - Hs) / Ht
gl.data$other$loc.metrics$Fst     <- (gl.data$other$loc.metrics$Ht - 
                                      gl.data$other$loc.metrics$meanHs) / 
                                      gl.data$other$loc.metrics$Ht
```

### Plot statistics

```{r popgen-plots}
par(mfrow = c(2, 4))

hist(
  gl.data$other$loc.metrics$mean_p,
  xlab = "Allele Freq (mean within pops)", 
  main = NULL, 
  ylab = NULL)

hist(
  gl.data$other$loc.metrics$meanHo,
  xlab = "Ho (mean within pops)",
  main = NULL, 
  ylab = NULL)

hist(
  gl.data$other$loc.metrics$meanHs,
  xlab = "Mean Hs", 
  main = NULL, 
  ylab = NULL)

hist(
  gl.data$other$loc.metrics$Ht,
  xlab = "Ht", 
  main = NULL, 
  ylab = NULL)

plot(
  gl.data$other$loc.metrics$meanHs, 
  gl.data$other$loc.metrics$meanHo,
  cex = 0.5, 
  col = "blue", 
  xlab = "Mean Hs", 
  ylab = "Ho")

hist(
  gl.data$other$loc.metrics$meanFis,
  xlab = "Fis", 
  main = NULL, 
  ylab = NULL)

hist(
  gl.data$other$loc.metrics$Fit,
  xlab = "Fit", 
  main = NULL, 
  ylab = NULL)

hist(
  gl.data$other$loc.metrics$Fst,
  xlab = "Fst", 
  main = NULL, 
  ylab = NULL)
```

### Plot pop gen metrics by SNP quality

```{r popgen-vs-quality}
par(mfrow = c(2, 3))

plot(
  gl.data$other$loc.metrics$CallRate, 
  gl.data$other$loc.metrics$meanFis,
  cex  = 0.4, 
  col  = "green", 
  xlab = "Call Rate", 
  ylab = "Fis")

plot(
  gl.data$other$loc.metrics$RepAvg, 
  gl.data$other$loc.metrics$meanFis,
  cex  = 0.4, 
  col  = "pink", 
  xlab = "Repeatability", 
  ylab = "Fis")

plot(
  depth_bothalleles, 
  gl.data$other$loc.metrics$meanFis,
  cex  = 0.4, 
  col  = "blue", 
  xlab = "Depth", 
  ylab = "Fis")

plot(
  depth_bothalleles, 
  gl.data$other$loc.metrics$meanHo,
  cex  = 0.4, 
  col  = "orange", 
  xlab = "Depth", 
  ylab = "Ho")

plot(
  gl.data$other$loc.metrics$CallRate, 
  gl.data$other$loc.metrics$meanHo,
  cex  = 0.4, 
  col  = "red", 
  xlab = "Call Rate", 
  ylab = "Ho")

plot(
  depth_bothalleles, 
  gl.data$other$loc.metrics$CallRate,
  cex  = 0.4, 
  col  = "purple", 
  xlab = "Depth", 
  ylab = "Call Rate")
```

## Filtering

### PCoA

```{r pcoa-prefilt}
pcoa <- gl.pcoa(gl.data)
gl.pcoa.scree(pcoa)
gl.pcoa.plot(pcoa, gl.data)

# 3D plot; shape can be cube or sphere (tetrahedron does not work)
gl.pcoa.plot.3d(
  pcoa, 
  gl.data,
  title  = "PCoA",
  xaxis  = 1, 
  yaxis  = 2, 
  zaxis  = 3,
  shape  = "sphere", 
  radius = 2,
  legend = "topright")
```

### Based on loci

```{r filter-loci}
# Call rate threshold — base on frequency rug plot from earlier
gl.data.filt <- gl.filter.callrate(
  gl.data,
  method    = "loc",
  threshold = 0.95,
  mono.rm   = TRUE,
  plot.out  = TRUE)

# Repeatability threshold — refer to earlier plots
gl.data.filt <- gl.filter.reproducibility(
  gl.data.filt,
  threshold = 0.95,
  v         = 3)

# MAF filter; MAF can have a noticeable effect on population structure
# (see Linck and Battey, 2019; Brockett et al., 2022)
gl.data.filt <- gl.filter.maf(
  gl.data.filt, 
  threshold = 0.05, 
  v         = 3)

# Check for monomorphs after locus filtering
gl.filter.monomorphs(gl.data.filt)
```

### Based on individuals

```{r filter-inds}
ind_callrate <- gl.report.callrate(
  gl.data.filt, 
  method    = "ind", 
  v         = 3)

gl.data.filt <- gl.filter.callrate(
  gl.data.filt,
  method    = "ind",
  threshold = 0.90,
  mono.rm   = TRUE,
  recalc    = FALSE,
  plot.out  = TRUE,
  v         = 3)

# Check for monomorphs after individual filtering
gl.filter.monomorphs(gl.data.filt)
```

### Save DartR object

```{r save-filtered}
save(gl.data.filt, file = "gl_maf05_allinfo.R")
```

### Check rug plot

```{r rug-plot}
# Plot rug plot before and after filtering together
par(mfrow = c(2, 1))
glPlot(gl.data)
glPlot(gl.data.filt)
```

### Plot pop gen 

```{r popgen-postfilt}
# Recalculate pop gen metrics after individual removal
gl.data.filt$other$loc.metrics$alf <- gl.alf(gl.data.filt)[, 2]

allelefreq1 <- aggregate(
  gl.data.filt, list(gl.data.filt$pop),
  function(x) {(mean(x, na.rm = TRUE) / 2)})

hetcount <- aggregate(
  gl.data.filt, list(gl.data.filt$pop),
  function(x) {sum(x[x == "1"], na.rm = TRUE)})

samplsize <- aggregate(
  gl.data.filt, list(gl.data.filt$pop),
  function(x) {length(x[!is.na(x)])})
```

```{r}
Hs <- 1 - 
  (allelefreq1[, -1]^2) - 
  ((1 - allelefreq1[, -1])^2)

Ho <- hetcount[, -1] / 
  samplsize[, -1]

Fis <- (Hs - Ho) / Hs
gl.data.filt$other$loc.metrics$Ht <- 2 * 
  gl.data.filt$other$loc.metrics$alf *
  (1 - gl.data.filt$other$loc.metrics$alf)

gl.data.filt$other$loc.metrics$meanHs  <- colMeans(Hs,  na.rm = TRUE)
gl.data.filt$other$loc.metrics$meanHo  <- colMeans(Ho,  na.rm = TRUE)
gl.data.filt$other$loc.metrics$meanFis <- colMeans(Fis, na.rm = TRUE)
gl.data.filt$other$loc.metrics$Ho      <- colSums(
  hetcount[, -1], na.rm = TRUE) / colSums(samplsize[, -1])

gl.data.filt$other$loc.metrics$Fit <- (
  gl.data.filt$other$loc.metrics$Ht - 
    gl.data.filt$other$loc.metrics$Ho) / 
  gl.data.filt$other$loc.metrics$Ht

gl.data.filt$other$loc.metrics$mean_p <- colMeans(
  allelefreq1[-1], na.rm = TRUE)

gl.data.filt$other$loc.metrics$Fst <- (
  gl.data.filt$other$loc.metrics$Ht - 
    gl.data.filt$other$loc.metrics$meanHs) / 
  gl.data.filt$other$loc.metrics$Ht

depth_bothalleles <- gl.data.filt$other$loc.metrics$AvgCountRef +
  gl.data.filt$other$loc.metrics$AvgCountSnp

alleleDepthRatio <- gl.data.filt$other$loc.metrics$AvgCountRef /
  gl.data.filt$other$loc.metrics$AvgCountSnp
```

```{r}
# Plot pop gen stats for post-filtered dataset
par(mfrow = c(2, 4))

hist(
  gl.data.filt$other$loc.metrics$mean_p,
  xlab = "Allele Freq - mean within pops", 
  main = NULL, 
  ylab = NULL)

hist(
  gl.data.filt$other$loc.metrics$meanHo,
  xlab = "Mean Ho", 
  main = NULL, 
  ylab = NULL)

hist(
  gl.data.filt$other$loc.metrics$meanHs,
  xlab = "Mean Hs", 
  main = NULL, 
  ylab = NULL)

hist(
  gl.data.filt$other$loc.metrics$Ht,
  xlab = "Ht", 
  main = NULL, 
  ylab = NULL)

plot(
  gl.data.filt$other$loc.metrics$meanHs, 
  gl.data.filt$other$loc.metrics$Ho,
  cex  = 0.5, 
  col  = "blue", 
  xlab = "Mean Hs", 
  ylab = "Ho")

hist(
  gl.data.filt$other$loc.metrics$meanFis,
  xlab = "Fis", 
  main = NULL, 
  ylab = NULL)

hist(
  gl.data.filt$other$loc.metrics$Fit,
  xlab = "Fit", 
  main = NULL, 
  ylab = NULL)

hist(
  gl.data.filt$other$loc.metrics$Fst,
  xlab = "Fst", 
  main = NULL, 
  ylab = NULL)
```

```{r}
# Plot basic SNP quality metrics for post-filtered dataset
par(mfrow = c(2, 2))

hist(
  gl.data.filt$other$loc.metrics$CallRate,
  xlab = "Call Rate", 
  main = NULL)

hist(
  depth_bothalleles,
  xlab = "Depth", 
  main = NULL)

hist(
  alleleDepthRatio,
  xlab = "Allele Seq Depth Ratio", 
  main = NULL)

hist(
  gl.data.filt$other$loc.metrics$RepAvg,
  xlab = "Repeatability", 
  main = NULL)
```

### Summarise genetic diversity

```{r popgen-summary}
postfilt_Fst    <- mean(gl.data.filt$other$loc.metrics$Fst,     na.rm = TRUE)
postfilt_Ho     <- mean(gl.data.filt$other$loc.metrics$meanHo,  na.rm = TRUE)
postfilt_Hs     <- mean(gl.data.filt$other$loc.metrics$meanHs,  na.rm = TRUE)
postfilt_Ht     <- mean(gl.data.filt$other$loc.metrics$Ht,      na.rm = TRUE)
postfilt_Fis    <- mean(gl.data.filt$other$loc.metrics$meanFis, na.rm = TRUE)

popgen_postfilt <- c(
  postfilt_Hs, 
  postfilt_Ho, 
  postfilt_Ht,
  postfilt_Fis, 
  postfilt_Fst)
```

```{r}
# Compare to pre-filtering values
prefilt_Fst    <- mean(gl.data$other$loc.metrics$Fst,     na.rm = TRUE)
prefilt_Ho     <- mean(gl.data$other$loc.metrics$meanHo,  na.rm = TRUE)
prefilt_Hs     <- mean(gl.data$other$loc.metrics$meanHs,  na.rm = TRUE)
prefilt_Ht     <- mean(gl.data$other$loc.metrics$Ht,      na.rm = TRUE)
prefilt_Fis    <- mean(gl.data$other$loc.metrics$meanFis, na.rm = TRUE)

popgen_prefilt <- c(
  prefilt_Hs, 
  prefilt_Ho, 
  prefilt_Ht,
  prefilt_Fis, 
  prefilt_Fst)
```

```{r}
popgen_summary <- rbind(popgen_prefilt, popgen_postfilt)

colnames(popgen_summary) <- c("Hs", "Ho", "Ht", "Fis", "Fst")
popgen_summary

write.csv(popgen_summary, file = "popgen_summary_prepost_maf05.csv")
```

# **Reproductive trends**

```{r repro-setup}
repro_data <- read.csv("Parentage_over_time.csv")
```

## Mann-Kendall tests

```{r mann-kendall}
mk_offspring            <- mk.test(repro_data$offspring)
mk_pct_females          <- mk.test(repro_data$pct_females_mothers)
mk_offspring_per_mother <- mk.test(repro_data$avg_offspring_per_mother)
mk_pct_males            <- mk.test(repro_data$pct_males_fathers)
mk_offspring_per_father <- mk.test(repro_data$avg_offspring_per_father)

cat("=== MANN-KENDALL TREND TESTS ===\n")
mk_offspring
mk_pct_females
mk_offspring_per_mother
mk_pct_males
mk_offspring_per_father
```

## Linear regression

```{r linear-regression}
run_regression <- function(y_var, metric_name) {
  model <- lm(y_var ~ year, data = repro_data)
  model_summary <- summary(model)

  cat("\nLinear Regression:", metric_name, "vs Year\n")
  cat("========================\n")
  cat("Slope:", model_summary$coefficients[2, 1], "per year\n")
  cat("p-value:", model_summary$coefficients[2, 4], "\n")
  cat("R-squared:", model_summary$r.squared, "\n")
  cat("Interpretation:",
      ifelse(model_summary$coefficients[2, 4] < 0.05,
             "Significant trend",
             ifelse(model_summary$coefficients[2, 4] < 0.1,
                    "Marginally significant trend",
                    "No significant trend")), "\n\n")

  return(model)
}
```

```{r}
cat("\n=== LINEAR REGRESSION ANALYSIS ===\n")

reg_offspring <- run_regression(
  repro_data$offspring, 
  "Total Offspring")

reg_pct_females <- run_regression(
  repro_data$pct_females_mothers,
  "% Females that are Mothers")

reg_offspring_per_mother <- run_regression(
  repro_data$avg_offspring_per_mother,
  "Avg Offspring per Mother")

reg_pct_males <- run_regression(
  repro_data$pct_males_fathers,
  "% Males that are Fathers")

reg_offspring_per_father <- run_regression(
  repro_data$avg_offspring_per_father,
  "Avg Offspring per Father")
```

```{r}
cat("\n=== SPEARMAN RANK CORRELATION ===\n")

sp_offspring <- run_spearman(
  repro_data$offspring, 
  "Total Offspring")

sp_pct_females <- run_spearman(
  repro_data$pct_females_mothers,
  "% Females that are Mothers")

sp_offspring_per_mother <- run_spearman(
  repro_data$avg_offspring_per_mother,
  "Avg Offspring per Mother")

sp_pct_males <- run_spearman(
  repro_data$pct_males_fathers,
  "% Males that are Fathers")

sp_offspring_per_father <- run_spearman(
  repro_data$avg_offspring_per_father,
  "Avg Offspring per Father")
```

## Chi-square test

```{r chi-square}
run_chi_square <- function(
    success1, total1, success2, total2, year1, year2, metric_name) {
  
  # Contingency table: rows = years, cols = success/failure counts
  contingency <- matrix(c(
    success1, 
    total1 - success1, 
    success2, 
    total2 - success2),
    nrow   = 2, 
    byrow  = TRUE)

  chi_result <- chisq.test(contingency, correct = TRUE)

  prop1 <- success1 / total1
  prop2 <- success2 / total2

  cat("\nChi-square Test:", metric_name, "from", year1, "to", year2, "\n")
  cat("========================\n")
  cat(year1, ":", round(prop1 * 100, 1), "% (", success1, "/", total1, ")\n",
      sep = "")
  cat(year2, ":", round(prop2 * 100, 1), "% (", success2, "/", total2, ")\n",
      sep = "")
  cat("Change:", round((prop2 - prop1) * 100, 1), "percentage points\n",
      sep = "")
  cat("Chi-square:", round(chi_result$statistic, 4), "\n")
  cat("p-value:", round(chi_result$p_value, 4), "\n")
  cat("Interpretation:",
      ifelse(chi_result$p.value < 0.05,
             "Significant difference",
             "No significant difference"), "\n\n")

  return(chi_result)
}
```

```{r}
# Female reproductive participation
cat("\n--- Percentage of Females that are Mothers ---\n")

female_2016_2017 <- run_chi_square(
  repro_data$mothers[1], 
  repro_data$adult_females[1],
  repro_data$mothers[2], 
  repro_data$adult_females[2],
  2016, 
  2017, 
  "% Females as Mothers")

female_2017_2018 <- run_chi_square(
  repro_data$mothers[2], 
  repro_data$adult_females[2],
  repro_data$mothers[3], 
  repro_data$adult_females[3],
  2017, 
  2018, 
  "% Females as Mothers")

female_2018_2019 <- run_chi_square(
  repro_data$mothers[3], 
  repro_data$adult_females[3],
  repro_data$mothers[4], 
  repro_data$adult_females[4],
  2018, 
  2019, 
  "% Females as Mothers")

female_2019_2020 <- run_chi_square(
  repro_data$mothers[4], 
  repro_data$adult_females[4],
  repro_data$mothers[5], 
  repro_data$adult_females[5],
  2019, 
  2020, 
  "% Females as Mothers")
```

```{r}
# Male reproductive participation
cat("\n--- Percentage of Males that are Fathers ---\n")

male_2016_2017 <- run_chi_square(
  repro_data$fathers[1], 
  repro_data$adult_males[1],
  repro_data$fathers[2], 
  repro_data$adult_males[2],
  2016, 
  2017, 
  "% Males as Fathers")

male_2017_2018 <- run_chi_square(
  repro_data$fathers[2], 
  repro_data$adult_males[2],
  repro_data$fathers[3], 
  repro_data$adult_males[3],
  2017, 
  2018, 
  "% Males as Fathers")

male_2018_2019 <- run_chi_square(
  repro_data$fathers[3], 
  repro_data$adult_males[3],
  repro_data$fathers[4], 
  repro_data$adult_males[4],
  2018, 
  2019, 
  "% Males as Fathers")

male_2019_2020 <- run_chi_square(
  repro_data$fathers[4], 
  repro_data$adult_males[4],
  repro_data$fathers[5], 
  repro_data$adult_males[5],
  2019, 
  2020, 
  "% Males as Fathers")
```

## Plots

### Test plots

```{r repro-plots}
ggplot(repro_data, aes(x = year)) +
  
  geom_line(aes(
    y     = pct_females_mothers * 100, 
    color = "Females"), 
    size  = 1.2) +
  
  geom_point(aes(
    y     = pct_females_mothers * 100, 
    color = "Females"), 
    size  = 2) +
  
  geom_line(aes(
    y     = pct_males_fathers * 100, 
    color = "Males"), 
    size  = 1.2) +
  
  geom_point(aes(
    y     = pct_males_fathers * 100, 
    color = "Males"), 
    size  = 2) +
  
  labs(
    title = "Reproductive Participation by Sex",
    x     = "Year",
    y     = "Percentage of adult population of given sex",
    color = "Sex") +
  
  scale_color_manual(values = c(
    "Females" = "plum2", 
    "Males"   = "skyblue3")) +
  
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
ggplot(repro_data, aes(x = year)) +
  
  geom_line(
    aes(y        = avg_offspring_per_mother, 
        color    = "Per Mother"), 
    size         = 1.2) +
  
  geom_point(
    aes(y        = avg_offspring_per_mother, 
        color    = "Per Mother"), 
    size         = 3) +
  
  geom_line(
    aes(y        = avg_offspring_per_father, 
        color    = "Per Father"), 
    size         = 1.2) +
  
  geom_point(
    aes(y        = avg_offspring_per_father, 
        color    = "Per Father"), 
    size         = 3) +
  
  labs(
    title        = "Per Capita Reproductive Success",
    subtitle     = "Average Offspring per Reproductive Individual",
    x            = "Year",
    y            = "Average Offspring",
    color        = "Metric") +
  
  scale_color_manual(values = c(
    "Per Mother" = "deeppink3",
    "Per Father" = "darkblue")) +
  
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
repro_data$pct_females_mothers_pct <- repro_data$pct_females_mothers * 100
repro_data$pct_males_fathers_pct   <- repro_data$pct_males_fathers   * 100

ggplot(repro_data, 
       aes(x = factor(year), 
           y = offspring)) +
  
  geom_bar(
    stat  = "identity", 
    fill  = "darkolivegreen4", 
    alpha = 0.7) +
  
  geom_text(aes(
    label = offspring), 
    vjust = -0.5) +
  
  labs(
    title = "Total Offspring Production by Year",
    x     = "Year",
    y     = "Number of Offspring") +
  
  theme_minimal()
```

### Dual-axis plot

```{r dual-axis-plot}
# Reshape to long format for breeding success
breeding_data <- repro_data %>%
  
  select(
    year, 
    pct_females_mothers, 
    pct_males_fathers) %>%
  
  pivot_longer(
    cols        = c(pct_females_mothers, pct_males_fathers),
    names_to    = "sex_metric",
    values_to   = "breeding_percentage") %>%
  
  mutate(
    sex         = ifelse(grepl("females", sex_metric), "Female", "Male"),
    metric_type = "Breeding Success")

# Reshape to long format for population sizes
population_data <- repro_data %>%
  
  select(
    year, 
    adult_females, 
    adult_males) %>%
  
  pivot_longer(
    cols      = c(adult_females, adult_males),
    names_to  = "sex_pop",
    values_to = "population_size") %>%
  
  mutate(
    sex         = ifelse(grepl("females", sex_pop), "Female", "Male"),
    metric_type = "Population Size")
```


```{r}
trends <- ggplot(repro_data) +
  
  geom_line(aes(
    x     = year, 
    y     = pct_females_mothers,
    color = "% females adults"), 
    size  = 1.2) +
  
  geom_point(aes(
    x     = year, 
    y     = pct_females_mothers,
    color = "% females adults"), 
    size  = 2) +
  
  geom_line(aes(
    x     = year, 
    y     = pct_males_fathers,
    color = "% males adults"), 
    size  = 1.2) +
  
  geom_point(aes(
    x     = year, 
    y     = pct_males_fathers,
    color = "% males adults"), size = 2) +
  
  geom_line(aes(
    x        = year, 
    y        = adult_females * 2,
    color    = "Number of adult females"),
    linetype = "dashed", 
    size     = 1.2) +
  
  geom_point(aes(
    x = year, 
    y = adult_females * 2,
    color = "Number of adult females"),
    shape = 15, 
    size = 3) +
  
  geom_line(aes(
    x        = year, 
    y        = adult_males * 2,
    color    = "Number of adult males"),
    linetype = "dashed", 
    size     = 1.2) +
  
  geom_point(aes(
    x     = year, 
    y     = adult_males * 2,
    color = "Number of adult males"),
    shape = 18, 
    size  = 3) +
  
  scale_y_continuous(
    name     = "Reproductive Success Rate (%)",
    limits   = c(0, 120),
    sec.axis = sec_axis(~. / 2, name = "Adult Population Size")) +
  
  scale_color_manual(
    name   = "Metric",
    values = c(
      "% females adults"        = "purple4",
      "% males adults"          = "deepskyblue1",
      "Number of adult females" = "maroon",
      "Number of adult males"   = "darkslategray3")) +
  
  scale_x_continuous(breaks = repro_data$year) +
  
  labs(
    title = "Reproductive Success and Population Size by Year",
    x     = "Year") +
  
  theme_minimal(base_size = 12) +
  theme(
    panel.background    = element_rect(fill  = "white", color = NA),
    plot.background     = element_rect(fill  = "white", color = NA),
    panel.grid.major    = element_line(color = "gray90", size = 0.5),
    panel.grid.minor    = element_line(color = "gray95", size = 0.25),
    axis.title.y        = element_text(color = "black", size = 12),
    axis.title.y.right  = element_text(color = "black", size = 12),
    axis.title.x        = element_text(size  = 12),
    axis.text           = element_text(size  = 12, color = "black"),
    plot.title          = element_text(hjust = 0.5, size = 12, color = "black"),
    legend.text         = element_text(size  = 12, color = "black"),
    legend.background   = element_rect(fill  = "white", color = NA),
    legend.position     = "bottom",
    legend.title        = element_blank())

trends
ggsave("Reproductive success trends per year.png", trends)
```

# **Preparing model input**

## Read data

```{r modelling-setup}
set.seed(1234)
dat <- read.csv(file = "Parentage_data_formattedforglmm.csv")
head(dat)
summary(dat)

# Replace string "NA" with true NA values
dat[dat == "NA"]        <- NA

# Set categorical variables as factors
dat$sex                 <- as.factor(dat$sex)
dat$morph               <- as.factor(dat$morph)
dat$cohort              <- as.factor(dat$cohort)
dat$cluster             <- as.factor(dat$cluster)
dat$year                <- as.factor(dat$year)
dat$id                  <- as.factor(dat$id)

# Set continuous variables as numeric
dat$breedingsuccess     <- as.numeric(dat$breedingsuccess)
dat$offspring           <- as.numeric(dat$offspring)
dat$age                 <- as.numeric(dat$age)
dat$weight              <- as.numeric(dat$weight)
dat$indhet              <- as.numeric(dat$indhet)

# Scale continuous variables for comparability
# Note: use unscaled values for plotting
dat$age_scaled          <- scale(dat$age)
dat$weight_scaled       <- scale(dat$weight)
dat$indhet_scaled       <- scale(dat$indhet)
dat$parentgensim_scaled <- scale(dat$parentgensim)
```

## Subset data

```{r data-subsetting}
# Remove rows with NAs in key variables
nona <- subset(dat,  weight       != "NA")
nona <- subset(nona, morph        != "NA")
nona <- subset(nona, age          != "NA")
nona <- subset(nona, year         != "NA")
nona <- subset(nona, sex          != "NA")
nona <- subset(nona, parentgensim != "NA")
nona <- subset(nona, indhet       != "NA")
nona <- subset(nona, cohort       != "NA")
nona <- subset(nona, cluster      != "NA")
```

Individual 88C7E is a potential outlier with reproductive success recorded at 200g, which may reflect human error, inaccurate parentage assignment, or incorrect field measurements.

```{r}
pruned  <- subset(nona,    id != "88C7E")

# Within-sex subsets
males   <- subset(pruned, sex != "F")
females <- subset(pruned, sex != "M")

# Restrict to successful breeders to reduce the influence of zero-offspring
# individuals in models of offspring count
success <- subset(pruned, breedingsuccess == 1)
```

## Breeding success

### Statistics

```{r summary-stats}
breeding_summary <- pruned %>%
  group_by(sex) %>%
  summarize(
    n             = n(),
    success_rate  = mean(breedingsuccess, na.rm = TRUE),
    avg_offspring = mean(offspring,       na.rm = TRUE))
print(breeding_summary)

sexbreed_plot <- ggplot(
  breeding_summary,
  aes(x    = sex, 
      y    = success_rate, 
      fill = sex)) +
  
  geom_bar(stat = "identity") +
  
  geom_text(aes(label = n), vjust = -0.5) +
  
  labs(
    x     = "Sex", 
    y     = "Breeding Success Rate",
    title = "Breeding Success by Sex") +
  
  ylim(0, 0.7) +
  
  theme_minimal() +
  theme(legend.position = "none")
sexbreed_plot
```

```{r}
breeding_by_age <- nona %>%
  group_by(age) %>%
  summarize(
    n            = n(),
    success_rate = mean(breedingsuccess, na.rm = TRUE))
print(breeding_by_age)

bredage_plot <- ggplot(
  breeding_by_age, 
  aes(x = age, 
      y = success_rate)) +
  
  geom_bar(stat = "identity", fill = "steelblue") +
  
  geom_text(aes(label = n), vjust = -0.5) +
  
  labs(
    x = "Age", 
    y = "Breeding Success Rate",
    title = "Breeding Success by Age") +
  
  ylim(0, 0.7) +
  
  theme_minimal()
bredage_plot
```

```{r}
breeding_by_age_sex <- nona %>%
  group_by(age, sex) %>%
  summarize(
    n            = n(),
    success_rate = mean(breedingsuccess, na.rm = TRUE))
print(breeding_by_age_sex)

agesex_plot <- ggplot(
  breeding_by_age_sex,
  aes(x    = age, 
      y    = success_rate, 
      fill = sex)) +
  
  geom_bar(
    stat     = "identity", 
    position = "dodge") +
  
  geom_text(
    aes(label = n), 
    vjust     = -0.5,
    position  = position_dodge(width = 0.9)) +
  
  labs(
    x     = "Age", 
    y     = "Breeding Success Rate",
    title = "Breeding Success by Age and Sex") +
  
  ylim(0, 0.7) +
  
  theme_minimal()
agesex_plot
```

```{r}
breeding_by_morph <- nona %>%
  filter(!is.na(morph)) %>%
  group_by(morph) %>%
  summarize(
    n            = n(),
    success_rate = mean(breedingsuccess, na.rm = TRUE))
print(breeding_by_morph)

morph_plot <- ggplot(
  breeding_by_morph,
  aes(x    = morph, 
      y    = success_rate, 
      fill = morph)) +
  
  geom_bar(stat = "identity") +
  
  geom_text(aes(label = n), vjust = -0.5) +
  
  labs(
    x     = "Color Morph", 
    y     = "Breeding Success Rate",
    title = "Breeding Success by Morph") +
  
  theme_minimal() +
  theme(legend.position = "none")

morph_plot
```

### Statistics

#### Weight

```{r dist-weight}
# Violin plot comparing weight distributions for breeders vs non-breeders
success_violin <- ggplot(
  pruned, 
  aes(x = factor(breedingsuccess), 
      y = weight)) +
  geom_violin(alpha  = 0.7) +
  geom_boxplot(width = 0.1) +
  xlab("Breeding Success") +
  ylab("Weight (kg)") +
  theme_bw()
success_violin
```

```{r}
success_by_weight <- ggplot(data = pruned) +
  aes(
    x = weight, 
    y = breedingsuccess) +
  geom_point(alpha = 0.5) +
  geom_smooth(
    method      = glm,
    method.args = list(family = binomial),
    se          = TRUE,
    colour      = "black",
    fill        = "forestgreen") +
  xlab("Weight (kg)") +
  ylab("")
success_by_weight
```

```{r}
offspring_by_weight <- ggplot(data = pruned) +
  aes(
    x = weight, 
    y = offspring) +
  geom_smooth(
    method      = glm,
    method.args = list(family = poisson),
    se          = TRUE,
    colour      = "black",
    fill        = "forestgreen") +
  xlab("Weight (kg)") +
  ylab("")
offspring_by_weight
```

#### Sex

```{r dist-sex}
offspring_by_sex <- ggplot(
  nona, 
  aes(x = sex, 
      y = offspring)) +
  geom_boxplot(
    outlier.fill   = "black", 
    outlier.shape  = 1) +
  stat_summary(fun = mean) +
  xlab("Sex") +
  ylab("Number of Offspring")
offspring_by_sex
```

#### Morph

```{r dist-morph}
offspring_by_morph <- ggplot(
  nona, 
  aes(x = morph, 
      y = offspring)) +
  geom_boxplot(
    outlier.fill   = "black", 
    outlier.shape  = 1) +
  stat_summary(fun = mean) +
  xlab("Morph") +
  ylab("")
offspring_by_morph

success_by_morph <- ggplot(
  nona, 
  aes(x = morph, 
      y = breedingsuccess)) +
  geom_violin(alpha  = 0.7) +
  geom_boxplot(width = 0.05) +
  xlab("Morph") +
  ylab("")
success_by_morph
```

```{r}
empirical_rates <- pruned %>%
  group_by(morph) %>%
  summarize(
    success_rate = mean(breedingsuccess, na.rm = TRUE),
    sample_size  = n(),
    se           = sqrt(success_rate * (1 - success_rate) / sample_size),
    lower        = pmax(0, success_rate - 1.96 * se),
    upper        = pmin(1, success_rate + 1.96 * se))

empirical_plot <- ggplot(
  empirical_rates,
  aes(
    x    = morph, 
    y    = success_rate, 
    fill = morph)) +
  
  geom_bar(stat = "identity", width = 0.6, alpha = 0.7) +
  
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  
  geom_text(aes(label = paste("n =", sample_size)), vjust = -0.8) +
  
  scale_fill_manual(values = c("darkgreen", "tan4")) +
  
  labs(
    x     = "Morph",
    y     = "Observed Breeding Success Rate",
    title = "Observed Breeding Success Rate by Morph") +
  
  ylim(0, max(empirical_rates$upper) * 1.1) +
  
  theme_minimal() +
  theme(
    legend.position  = "none",
    axis.line        = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    panel.border     = element_blank(),
    text             = element_text(size = 14, colour = "black"))
```

#### Age

```{r dist-age}
success_by_age_violin <- ggplot(
  nona,
  aes(x = factor(breedingsuccess), 
      y = age)) +
  geom_violin(alpha  = 0.7) +
  geom_boxplot(width = 0.1) +
  xlab("Breeding Success") +
  ylab("Age (years)") +
  theme_bw()
success_by_age_violin

success_by_age <- ggplot(data = nona) +
  aes(x = age, 
      y = breedingsuccess) +
    geom_smooth(
    method      = glm,
    method.args = list(family = "binomial"),
    se          = TRUE,
    colour      = "black",
    fill        = "darkgoldenrod1") + 
  xlab("Age (years)") +
  ylab("Likelihood of Breeding Success")
success_by_age

offspring_by_age <- ggplot(data = nona) +
  aes(x = age, 
      y = offspring) +
  geom_smooth(
    method      = glm,
    method.args = list(family = "poisson"),
    se          = TRUE,
    colour      = "black",
    fill        = "darkgoldenrod1") +
  xlab("Age (years)") +
  ylab("Predicted number of offspring")
offspring_by_age

success_by_age_pruned <- ggplot(data = pruned) +
  aes(x = age, 
      y = breedingsuccess) +
  geom_smooth(
    method      = glm,
    method.args = list(family = "binomial"),
    se          = TRUE,
    colour      = "black",
    fill        = "darkgoldenrod1") +
  xlab("Age (years)") +
  ylab("Likelihood of Breeding Success")
success_by_age_pruned

offspring_by_age_pruned <- ggplot(data = pruned) +
  aes(x = age, 
      y = offspring) +
  geom_smooth(
    method      = glm,
    method.args = list(family = "poisson"),
    se          = TRUE,
    colour      = "black",
    fill        = "darkgoldenrod1") +
  xlab("Age (years)") +
  ylab("")
offspring_by_age_pruned
```

#### Individual heterozygosity

```{r dist-het}
success_by_het_violin <- ggplot(
  nona,
  aes(x = factor(breedingsuccess), 
      y = indhet)) +
  geom_violin(alpha  = 0.7) +
  geom_boxplot(width = 0.1) +
  xlab("Breeding Success") +
  ylab("Individual Heterozygosity") +
  theme_bw()
success_by_het_violin

success_by_het <- ggplot(data = nona) +
  aes(x = indhet, 
      y = breedingsuccess) +
  geom_smooth(
    method      = glm,
    method.args = list("binomial"),
    se          = TRUE,
    colour      = "black",
    fill        = "firebrick3") +
  xlab("Individual Heterozygosity") +
  ylab("")
success_by_het

offspring_by_het <- ggplot(data = nona) +
  aes(x = indhet, 
      y = offspring) +
  geom_smooth(
    method      = glm,
    method.args = list("poisson"),
    se          = TRUE,
    colour      = "black",
    fill        = "firebrick4") +
  xlab("Individual Heterozygosity") +
  ylab("")
offspring_by_het
```

```{r trend-plots}
trends <- (success_by_weight | offspring_by_weight) /
  (success_by_age            | offspring_by_age)    /
  (success_by_morph          | offspring_by_morph)  /
  (success_by_het            | offspring_by_het)    & theme_classic()
trends
```

## Breeding success

### Global model

```{r binomial-model}
binmod_pruned <- glmmTMB(
  breedingsuccess ~
    (scale(weight) + I((scale(weight))^2)) * scale(age) +
    scale(indhet)  + sex + morph +
    (1 | id) +
    (1 | year),
  family    = "binomial",
  data      = pruned,
  na.action = na.fail)
```

```{r}
# Check global model fit before dredging (as per Mazerolle, 2020)
check_overdispersion(      binmod_pruned)
check_collinearity(        binmod_pruned)
ch_bin_nona <- check_model(binmod_pruned)
ch_bin_nona

summary(binmod_pruned)

plot_model(
  binmod_pruned, 
  type   = "pred",
  terms  = "weight[all]",
  ci.lvl = 0.95)
```

```{r}
# Test whether a subset of factors produces a better model
modsel_pruned     <- dredge(binmod_pruned, trace = 2)
modsel_sub_pruned <- subset(modsel_pruned, delta < 5)
modsel_sub_pruned
```

The top model has the lowest degrees of freedom and subsequent models offer minimal improvement in fit; select the top model rather than averaging. Weight has the greatest effect on breeding success (positive); age and fawn morph are both negatively associated with success. The slight negative quadratic for weight likely reflects age overriding weight benefits as individuals get heavier with age.

```{r}
topmod_pruned <- get.models(modsel_sub_pruned, subset = 1)[[1]]
topmod_pruned
summary(topmod_pruned)

check_overdispersion(topmod_pruned)
check_collinearity(topmod_pruned)
ch_bin_topmod_pruned <- check_model(topmod_pruned)

plot_model(topmod_pruned)
```

### Model table

```{r binomial-model-table}
modeltabletest <- tab_model(topmod_pruned)
model_table <- tab_model(
  topmod_pruned,
  transform   = "exp",
  show.est    = TRUE,
  show.stat   = TRUE,
  show.p      = TRUE,
  show.ci     = 0.95,
  show.re.var = TRUE,
  show.aic    = FALSE,
  show.r2     = TRUE,
  show.icc    = FALSE,
  collapse.ci = TRUE,
  title       = "Factors Affecting Breeding Success",
  dv.labels   = "Predicted impact",
  pred.labels = c(
    "(Intercept)"          = "Intercept",
    "scale(weight)"        = "Weight (kg)",
    "I((scale(weight))^2)" = "Weight (kg) quadratic",
    "scale(age)"           = "Age (years)",
    "morphFawn"            = "Morph (fawn)"),
  string.est  = "Odds Ratios (95% CI)",
  string.stat = "z-statistic",
  string.p    = "p-value",
  file        = "breeding_success_model.html")

model_table
webshot(
  "breeding_success_model.html",
  "breeding_success_model.png",
  vwidth  = 800, 
  vheight = 600)
```

### Predictor plots

#### Weight

```{r plot-weight-breeding}
summary(topmod_pruned)

# Refit top model explicitly for clean prediction (no interaction term)
topmod_pruned <- glmmTMB(
  breedingsuccess ~
    scale(weight) + I((scale(weight))^2) + scale(age) + morph + (1 | id),
  family    = "binomial",
  data      = pruned,
  na.action = na.fail)

# Prediction grid: vary weight, hold age and morph at reference values
pred_data <- data.frame(
  weight = seq(
    min(pruned$weight, na.rm = TRUE),
    max(pruned$weight, na.rm = TRUE),
    length.out = 100),
  age   = mean(pruned$age, na.rm = TRUE),
  morph = levels(pruned$morph)[1],
  id    = NA)

preds <- predict(
  topmod_pruned, 
  newdata = pred_data, 
  se.fit  = TRUE,
  re.form = NA)
```

```{r}
# Assemble predictions on link scale, then convert to probability
plotdat <- data.frame(
  weight   = pred_data$weight,
  fit_link = preds$fit,
  se_link  = preds$se.fit) %>%
  mutate(
    lower_link     = fit_link - 1.96 * se_link,
    upper_link     = fit_link + 1.96 * se_link,
    fit_response   = plogis(fit_link),
    lower_response = plogis(lower_link),
    upper_response = plogis(upper_link),
    
    # Weight at which predicted breeding success is highest
    bestweight     = weight[which.max(fit_response)])
```

```{r}
plotweightbreed <- {ggplot(
  data = plotdat,
  aes(
    x    = weight, 
    y    = fit_response,
    ymin = lower_response, 
    ymax = upper_response,
    xmin = min(pruned$weight), 
    xmax = max(pruned$weight))) +
    
  geom_ribbon(
    alpha     = 0.2,
    colour    = "darkgreen",
    fill      = "darkgreen",
    linetype  = "dotted") +
    
  geom_line(
    linetype  = "solid",
    colour    = "green4",
    linewidth = 1) +
    
  xlab("Weight (kg)") +
  ylab("Likelihood of Breeding Success") +
  theme(
    axis.line        = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    panel.border     = element_blank(),
    axis.text.y      = element_text(angle = 0, vjust = 0.5, color = "black"),
    axis.text.x      = element_text(color = "black"),
    strip.text.x     = element_text(colour = "black"),
    plot.margin      = margin(0.5, 0.5, 0.5, 0.5, "cm"),
    text             = element_text(
      size = 14, colour = "black", face = "plain"))}

plotweightbreed
ggsave(
  "breeding_success_by_weight.png", 
  plotweightbreed,
  width  = 8, 
  height = 6, 
  dpi    = 300)
```

#### Age

```{r plot-age-breeding}
# Scale age using training data mean and sd
age_mean   <- mean(pruned$age, na.rm = TRUE)
age_sd     <- sd(pruned$age, na.rm = TRUE)
age_values <- seq(
  min(pruned$age, na.rm = TRUE),
  max(pruned$age, na.rm = TRUE),
  length.out = 100)

# Extract coefficients needed to construct predictions manually
age_coef           <- simple_coefs["scale(age)",             "Estimate"]
age_se             <- simple_coefs["scale(age)",             "Std. Error"]
intercept          <- simple_coefs["(Intercept)",            "Estimate"]
weight_coef        <- simple_coefs["scale(weight)",          "Estimate"]
weight_sq_coef     <- simple_coefs["I((scale(weight))^2)",   "Estimate"]
weight_mean_scaled <- mean(scale(pruned$weight), na.rm = TRUE)
weight_sq_mean     <- mean((scale(pruned$weight))^2, na.rm = TRUE)
```

```{r}
# Manual prediction: linear predictor across age range
manual_pred <- data.frame(
  age        = age_values,
  age_scaled = (age_values - age_mean) / age_sd)

manual_pred$linear_pred <- intercept +
  weight_coef    * weight_mean_scaled +
  weight_sq_coef * weight_sq_mean +
  age_coef       * manual_pred$age_scaled

manual_pred$probability <- plogis(manual_pred$linear_pred)

# 95% CI on link scale, convert to probability
manual_pred$lower_linear <- manual_pred$linear_pred -
  1.96 * age_se * abs(manual_pred$age_scaled)

manual_pred$upper_linear <- manual_pred$linear_pred +
  1.96 * age_se * abs(manual_pred$age_scaled)

manual_pred$lower_prob <- plogis(manual_pred$lower_linear)
manual_pred$upper_prob <- plogis(manual_pred$upper_linear)

best_age <- manual_pred$age[which.max(manual_pred$probability)]
```

```{r}
age_plot <- {ggplot(manual_pred, aes(x = age, y = probability)) +
  
  geom_ribbon(aes(
    ymin = lower_prob, 
    ymax = upper_prob),
    alpha = 0.2,
    fill  = "darkblue") +
  
  geom_line(color = "blue4", size = 1) +
  
  labs(
    x     = "Age (years)",
    y     = "Likelihood of Breeding Success",
    title = "Effect of Age on Breeding Success (manual calculation)") +
  
  theme(
    axis.line        = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    panel.border     = element_blank(),
    axis.text.y      = element_text(angle = 0, vjust = 0.5, color = "black"),
    axis.text.x      = element_text(color = "black"),
    strip.text.x     = element_text(colour = "black"),
    plot.margin      = margin(0.5, 0.5, 0.5, 0.5, "cm"),
    text             = element_text(
      size = 14, colour = "black", face = "plain"))}

age_plot
ggsave(
  "age_effect_on_breeding_success.png",
  plot   = age_plot, 
  width  = 8, 
  height = 6, 
  dpi    = 300)
```

#### Morph

```{r plot-morph-breeding}
morph_violin <- ggplot(
  pruned, 
  aes(
    x    = morph, 
    y    = breedingsuccess, 
    fill = morph)) +
  
  geom_violin(alpha = 0.6) +
  geom_boxplot(width = 0.07, alpha = 0.7) +
  
  geom_jitter(
    width  = 0.15, 
    height = 0.02, 
    alpha  = 0.5, 
    colour = "black") +
  
  scale_fill_manual(values = c("grey30", "tan4")) +
  
  labs(
    x     = "Morph",
    y     = "Breeding Success",
    title = "Raw Breeding Success by Morph") +
  
  theme_minimal() +
  theme(
    legend.position  = "none",
    axis.line        = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    panel.border     = element_blank(),
    text             = element_text(size = 14, colour = "black"))
morph_violin
```

Morph predictions are constructed manually from model coefficients. All continuous predictors are held at their scaled mean (= 0). Standard errors are propagated separately for the reference morph (intercept SE only) and the fawn morph (combined intercept and morph coefficient SE).

```{r}
summary(topmod_pruned)$coefficients

# Hardcoded coefficients from model summary for manual prediction
morph_coef <- -2.4320712
morph_se   <-  0.8101449
morph_p    <- -3.002020
intercept  <-  1.1873517

# Hold other predictors at their scaled mean (= 0 after scaling)
weight_scaled <- 0
age_scaled    <- 0

base_linear_pred <- intercept +
  ( 2.5050649 * weight_scaled) +
  (-0.6925422 * weight_scaled^2) +
  (-1.1900311 * age_scaled)
```

```{r}
morph_levels <- levels(pruned$morph)
pred_data    <- data.frame(morph = factor(morph_levels, levels = morph_levels))

# Apply morph coefficient to fawn category only
pred_data$linear_pred             <- base_linear_pred
fawn_idx                          <- which(pred_data$morph == "Fawn")
pred_data$linear_pred[fawn_idx]   <- base_linear_pred + morph_coef
pred_data$predicted_success       <- plogis(pred_data$linear_pred)

# SE propagation: reference morph uses intercept SE only;
# fawn morph uses combined intercept + morph SE
pred_data$se_link                              <- 0
pred_data$se_link[pred_data$morph == levels(pruned$morph)[1]] <- 
  sqrt(0.5862252^2)
pred_data$se_link[pred_data$morph == "Fawn"]  <- 
  sqrt(0.5862252^2 + 0.8101449^2)

pred_data$lower_link <- pred_data$linear_pred - 1.96 * pred_data$se_link
pred_data$upper_link <- pred_data$linear_pred + 1.96 * pred_data$se_link
pred_data$lower      <- plogis(pred_data$lower_link)
pred_data$upper      <- plogis(pred_data$upper_link)
```

```{r}
morph_plot <- ggplot(
  pred_data,
  aes(
    x    = morph, 
    y    = predicted_success, 
    fill = morph)) +
  
  geom_bar(stat = "identity", width = 0.6, alpha = 0.7) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  scale_fill_manual(values = c("black", "tan3")) +
  
  labs(
    x     = "Morph",
    y     = "Predicted Breeding Success",
    title = "Predicted Breeding Success by Morph") +
  
  theme_minimal() +
  theme(
    legend.position  = "none",
    axis.line        = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    panel.border     = element_blank(),
    text             = element_text(size = 14, colour = "black"))

morph_plot
ggsave(
  "predicted_success_by_morph_manual.png", 
  morph_plot,
  width  = 8, 
  height = 6, 
  dpi    = 300)
```

#### Combined

```{r combined-figure}
morph_plot2 <- {ggplot(
  pred_data,
  aes(
    x    = morph, 
    y    = predicted_success, 
    fill = morph)) +
    
  geom_bar(
    stat  = "identity", 
    width = 0.6, 
    alpha = 0.8) +
  
  geom_errorbar(aes(
    ymin  = lower, 
    ymax  = upper),
    width = 0.2, 
    color = "black") +
  
  scale_fill_manual(values = c("black", "tan3")) +
  
  labs(
    x     = "Morph", 
    y     = NULL, 
    title = "C)") +
  
  ylim(0, 1) +
  
  theme_minimal() +
  theme(
    legend.position      = "none",
    axis.line            = element_line(colour = "black"),
    panel.grid.major.y   = element_blank(),
    panel.grid.minor     = element_blank(),
    panel.grid.major.x   = element_line(color = "gray90"),
    panel.background     = element_blank(),
    panel.border         = element_blank(),
    plot.title           = element_text(face = "bold", size = 12),
    axis.title.y         = element_text(size = 10),
    axis.text            = element_text(color = "black", size = 9))}
```

```{r}
age_plot2 <- {ggplot(
  manual_pred, aes(
    x = age, 
    y = probability)) +
  
  geom_ribbon(aes(
    ymin  = lower_prob, 
    ymax  = upper_prob),
    fill  = "blue", 
    alpha = 0.2) +
  
  geom_line(color = "blue4", size = 1) +
  
  labs(
    x     = "Age (years)", 
    y     = NULL, 
    title = "A)") +
  
  ylim(0, 1) +
  
  theme_minimal() +
  theme(
    legend.position    = "none",
    axis.line          = element_line(colour = "black"),
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_line(color = "gray90"),
    panel.background   = element_blank(),
    panel.border       = element_blank(),
    plot.title         = element_text(face = "bold", size = 12),
    axis.title.y       = element_text(size = 10),
    axis.text          = element_text(color = "black", size = 9))}
```

```{r}
weight_plot2 <- {ggplot(
  plotdat, aes(
    x = weight, 
    y = fit_response)) +
  
  geom_ribbon(aes(
    ymin  = lower_response, 
    ymax  = upper_response),
    fill  = "darkgreen", 
    alpha = 0.2) +
  
  geom_line(color = "green4", size = 1) +
  
  labs(
    x     = "Weight (kilograms)", 
    y     = NULL, 
    title = "B)") +
  
  ylim(0, 1) +
  
  theme_minimal() +
  theme(
    legend.position    = "none",
    axis.line          = element_line(colour = "black"),
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_line(color = "gray90"),
    panel.background   = element_blank(),
    panel.border       = element_blank(),
    plot.title         = element_text(face = "bold", size = 12),
    axis.title.y       = element_text(size = 10),
    axis.text          = element_text(color = "black", size = 9))}
```

```{r}
combined_figure <- ((age_plot2 / weight_plot2) | morph_plot2) +
  plot_layout(ncol = 2, widths = c(1.5, 0.7)) +
  plot_annotation(
    title = "Factors Affecting Breeding Success",
    theme = theme(plot.title = element_text(
      size  = 16, 
      face  = "bold", 
      hjust = 0.5)))
combined_figure

# Add shared y-axis label via grid
combined_grob <- patchworkGrob(combined_figure)

y_label <- textGrob(
  "Predicted likelihood of breeding success",
  rot   = 90,
  gp    = gpar(fontsize = 14),
  vjust = 1)

final_figure <- grid.arrange(
  y_label, 
  combined_grob,
  ncol   = 2,
  widths = c(0.05, 0.95))

ggsave(
  "Factors_affecting_breeding_success.png", 
  final_figure,
  width  = 8, 
  height = 6,
  dpi    = 300)
```

# **Offspring count**

## Global model

This is a pseudo-hurdle model. The first stage modelled drivers of reproductive success or failure (binomial above). This second stage models, for successfully reproducing individuals only, drivers of higher fecundity.

```{r poisson-model}
summary(success)
# 31 observations across 4 years; 24 unique individuals (12 M, 12 F)

psn_success <- glmmTMB(
  offspring ~
    I((scale(weight))^2) * scale(age) +
    scale(weight) +
    scale(age) +
    scale(indhet) +
    morph +
    sex +
    (1 | id) +
    (1 | year),
  data      = success,
  family    = "poisson",
  na.action = na.fail)

check_overdispersion(psn_success)
check_collinearity(psn_success)
check_model(psn_success)
summary(psn_success)
```

```{r}
# Fit nbinom2 as a comparison
nb_success <- glmmTMB(
  offspring ~
    I((scale(weight))^2) * scale(age) +
    scale(weight) +
    scale(age) +
    scale(indhet) +
    morph +
    sex +
    (1 | id) +
    (1 | year),
  data = success,
  family = "nbinom2")

check_overdispersion(nb_success)
check_model(nb_success)

# Poisson model has lower AIC and one fewer df; more parsimonious
AIC(psn_success, nb_success)
```

```{r}
# Dredge to check whether any subset of predictors improves fit
psnsel     <- dredge(psn_success)
psnsel_sub <- subset(psnsel, delta < 5)
psnsel_sub

write.csv(psnsel_sub, "model_selection_table.csv", row.names = FALSE)
```

```{r}
# Select top model (lowest df)
topmod_psn <- get.models(psnsel_sub, subset = 1)[[1]]
topmod_psn
summary(topmod_psn)

check_overdispersion(topmod_psn)
check_collinearity(topmod_psn)
check_model(topmod_psn)
```

## Model table

```{r poisson-model-table}
model_table_offspring <- tab_model(
  topmod_psn,
  transform   = "exp",
  show.est    = TRUE,
  show.stat   = TRUE,
  show.p      = TRUE,
  show.ci     = 0.95,
  show.re.var = TRUE,
  show.aic    = FALSE,
  show.r2     = TRUE,
  show.icc    = FALSE,
  collapse.ci = TRUE,
  pred.labels = c(
    "(Intercept)" = "Intercept",
    "scale(age)"  = "Age (years)",
    "morphFawn"   = "Morph (fawn)"),
  dv.labels   = "Predicted impact",
  string.est  = "Odds Ratios (95% CI)",
  string.stat = "z-statistic",
  string.p    = "p-value",
  file        = "offspring_model.html")

model_table_offspring

webshot(
  "offspring_model.html",
  "offspring_model.png",
  vwidth  = 800, 
  vheight = 600)
```

## Sex-specific models

### Modelling females

#### Binomial

##### Model

```{r female-binomial}
summary(females)

binmod_f <- glmmTMB(
  breedingsuccess ~
    scale(weight) +
    scale(age) +
    scale(indhet) +
    morph +
    (1 | id) + (1 | year),
  family    = "binomial",
  data      = females,
  na.action = na.fail)

check_model(binmod_f)
summary(binmod_f)
plot_model(binmod_f, type = "pred", ci.lvl = 0.95)
```

```{r}
modsel_f <- dredge(binmod_f)
modsel_f <- subset(modsel_f, delta < 5)
modsel_f

# Multiple models with low delta; average across models within delta < 2
binmod_f_avg <- model.avg(subset(modsel_f, delta < 2), fit = TRUE)
summary(binmod_f_avg)

# Diagnostic checks on full model (checks cannot be run on averaged models)
check_overdispersion(binmod_f)
check_collinearity(binmod_f)
check_model(binmod_f)
```

##### Plot

```{r plot-female-binomial}
weight_mean    <- mean(females$weight, na.rm = TRUE)
weight_sd      <- sd(  females$weight, na.rm = TRUE)
weight_kg_min  <- min( females$weight, na.rm = TRUE)
weight_kg_max  <- max( females$weight, na.rm = TRUE)

weight_kg_seq  <- seq(
  from          = weight_kg_min, 
  to            = weight_kg_max, 
  length.out    = 100)

weight_std_seq <- (weight_kg_seq - weight_mean) / weight_sd

# Hardcoded coefficients from model summary
intercept      <- -0.8555
weight_coef    <- 1.4213
intercept_se   <- 0.5955
weight_se      <- 0.5121

pred_link      <- intercept + weight_coef * weight_std_seq
z_68           <- qnorm(0.84)

# SE on link scale via error propagation
pred_se        <- sqrt(intercept_se^2 + (weight_se * weight_std_seq)^2)
lower_link     <- pred_link - z_68 * pred_se
upper_link     <- pred_link + z_68 * pred_se
pred_response  <- plogis(pred_link)
lower_response <- plogis(lower_link)
upper_response <- plogis(upper_link)

plotdat_f <- data.frame(
  weight_kg     = weight_kg_seq,
  weight_std    = weight_std_seq,
  fit           = pred_response,
  lower         = lower_response,
  upper         = upper_response)
```

```{r}
f_weight_plot <- ggplot(
  plotdat_f, 
  aes(
    x = weight_kg, 
    y = fit)) +
  
  geom_line(size = 1) +
  
  geom_ribbon(aes(
    ymin = lower, 
    ymax = upper),
    alpha = 0.2,
    fill  = "tomato3") +
  
  labs(
    title = "Effect of weight on female breeding success",
    x     = "Weight (kilograms)",
    y     = "Probability of Breeding Success") +
  
  theme_classic() +
  theme(
    plot.title  = element_text(size = 14, face = "bold"),
    axis.title  = element_text(size = 12),
    axis.text   = element_text(size = 10))
f_weight_plot
```

##### Model tables

```{r female-model-table}
model_table_f <- tab_model(
  binmod_f_avg,
  transform   = "exp",
  show.est    = TRUE,
  show.stat   = TRUE,
  show.p      = TRUE,
  show.ci     = 0.95,
  show.re.var = TRUE,
  show.aic    = FALSE,
  show.r2     = TRUE,
  show.icc    = FALSE,
  collapse.ci = TRUE,
  title       = "Factors Affecting Female Breeding Success",
  dv.labels   = "Predicted impact",
  pred.labels = c(
    "cond((Int))"          = "Intercept",
    "cond(morphFawn)"      = "Morph (fawn)",
    "cond(scale(age))"     = "Age (years)",
    "cond(scale(weight))"  = "Weight (kg)",
    "cond(scale(indhet))"  = "Heterozygosity"),
  string.est  = "Odds Ratios (95% CI)",
  string.stat = "z-statistic",
  string.p    = "p-value",
  file        = "fems_model.html")

model_table_f

webshot(
  "fems_model.html",
  "fems_model.png",
  vwidth  = 800, 
  vheight = 600)
```

```{r}
topmod_f <- get.models(modsel_sub_f, subset = 1)[[1]]
VarCorr(topmod_f)
```

#### Poisson

##### Model

```{r female-poisson}
# 13 successful females — very small dataset; model is exploratory only
fems <- subset(females, females$breedingsuccess == 1)
summary(fems)

mod_fems <- glmmTMB(
  offspring ~
    scale(weight) +
    scale(age) +
    scale(indhet) +
    morph +
    (1 | id) +
    (1 | year),
  family    = "poisson",
  data      = fems,
  na.action = na.fail)

summary(mod_fems)
check_overdispersion(mod_fems)
check_collinearity(mod_fems)
ch_f <- check_model(mod_fems)
ch_f
```

Too few observations to draw firm conclusions; predicted values not tracking observations well. 

```{r}
# Zero-inflated model on full female dataset; better residual fit
mod_f <- glmmTMB(
  offspring ~
    scale(weight) +
    scale(age) +
    morph +
    (1 | id) +
    (1 | year),
  family     = "poisson",
  data       = females,
  ziformula  = ~.,
  na.action  = na.fail)

summary(mod_f)
check_model(mod_f)
```

```{r}
modself      <- dredge(mod_f)
modsel_sub_f <- subset(modself, delta < 5)
modsel_sub_f

# Multiple models with low delta; use model averaging
topmod_f <- get.models(modsel_sub_f, subset = 1)[[1]]
summary(topmod_f)
```

No effects on female offspring count beyond those captured by the binomial.

### Modelling males

#### Binomial

#### Model

```{r male-binomial}
summary(males)

# 18 successful males
mals <- subset(males, males$breedingsuccess == 1)
summary(mals)

binmod_m <- glmmTMB(
  breedingsuccess ~
    scale(weight) +
    scale(age) +
    morph +
    (1 | id) + (1 | year),
  family    = "binomial",
  data      = males,
  na.action = na.fail)

check_model(binmod_m)
summary(binmod_m)

plot_model(binmod_m, type = "pred", ci.lvl = 0.95)
```

```{r}
modsel_m <- dredge(binmod_m)
modsel_m <- subset(modsel_m, delta < 5)
modsel_m

topmod_m <- get.models(modsel_m, subset = 1)[[1]]
summary(topmod_m)

check_overdispersion(topmod_m)
check_collinearity(topmod_m)
```

No significant results, though weight has strong support for affecting breeding success. 

#### Plot

```{r plot-male-binomial}
# Prediction grid across male weight range
pred_data_m <- data.frame(
  weight = seq(
    min(males$weight, na.rm = TRUE),
    max(males$weight, na.rm = TRUE),
    length.out = 100),
  age   = mean(males$age, na.rm = TRUE),
  morph = levels(males$morph)[1],
  id    = NA)

preds_m <- predict(
  topmod_m, 
  newdata = pred_data_m, 
  se.fit  = TRUE,
  re.form = NA)

# Assemble predictions and compute CI on probability scale
plotdat_m <- data.frame(
  weight_m           = pred_data_m$weight,
  fit_link_m         = preds_m$fit,
  se_link_m          = preds_m$se.fit) %>%
  mutate(
    lower_link_m     = fit_link_m - 1.96 * se_link_m,
    upper_link_m     = fit_link_m + 1.96 * se_link_m,
    fit_response_m   = plogis(fit_link_m),
    lower_response_m = plogis(lower_link_m),
    upper_response_m = plogis(upper_link_m),
    bestweight_m     = weight_m[which.max(fit_response_m)])
```

```{r}
plotbreed_m <- {ggplot(
  data = plotdat_m,
  aes(
    x    = weight_m, 
    y    = fit_response_m,
    ymin = lower_response_m, 
    ymax = upper_response_m,
    xmin = min(males$weight), 
    xmax = max(males$weight))) +
    
  geom_ribbon(
    alpha    = 0.2,
    colour   = "darkgreen",
    fill     = "darkgoldenrod",
    linetype = "dotted") +
    
  geom_line(
    linetype  = "solid",
    colour    = "black",
    linewidth = 1) +
    
  xlab("Weight (kg)") +
  ylab("Likelihood of Breeding Success") +
  theme(
    axis.line        = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    panel.border     = element_blank(),
    axis.text.y      = element_text(angle = 0, vjust = 0.5, color = "black"),
    axis.text.x      = element_text(color = "black"),
    strip.text.x     = element_text(colour = "black"),
    plot.margin      = margin(0.5, 0.5, 0.5, 0.5, "cm"),
    text             = element_text(size = 14, colour = "black", face = "plain"))}

plotbreed_m
ggsave(
  "breeding_success_males.png", 
  plotbreed_m,
  width  = 8, 
  height = 6, 
  dpi    = 300)
```

#### Model tables

```{r male-model-table}
summary(topmod_m)

model_table_m <- tab_model(
  topmod_m,
  transform   = "exp",
  show.est    = TRUE,
  show.stat   = TRUE,
  show.p      = TRUE,
  show.ci     = 0.95,
  show.re.var = TRUE,
  show.aic    = FALSE,
  show.r2     = TRUE,
  show.icc    = FALSE,
  collapse.ci = TRUE,
  title       = "Factors Affecting Male Breeding Success",
  dv.labels   = "Predicted impact",
  pred.labels = c(
    "Intercept"      = "Intercept",
    "morphFawn"      = "Morph (fawn)",
    "scale(age)"     = "Age (years)",
    "scale(weight)"  = "Weight (kg)"),
  string.est  = "Odds Ratios (95% CI)",
  string.stat = "z-statistic",
  string.p    = "p-value",
  file        = "males_model.html")

model_table_m

webshot(
  "males_model.html",
  "males_model.png",
  vwidth  = 800, 
  vheight = 600)
```

#### Poisson

```{r male-poisson}
mod_mals <- glmmTMB(
  offspring ~
    scale(weight) +
    scale(age) +
    scale(indhet) +
    morph +
    (1 | id) +
    (1 | year),
  family    = "poisson",
  data      = mals,
  na.action = na.fail)

summary(mod_mals)
ch_m <- check_model(mod_mals)
ch_m
```

Too few observations; predictions not aligning with data. 

```{r}
# Zero-inflated model on full male dataset; better residual fit
mod_m <- glmmTMB(
  offspring  ~ scale(age) + morph + (1 | id) + (1 | year),
  family     = "poisson",
  data       = males,
  ziformula  = ~.,
  na.action  = na.fail)

summary(mod_m)
check_model(mod_m)
```

No effects of offspring count in males; may reflect small dataset. Potentially indicates that population-level effects of morph, weight, and age affect female reproductive output more than male.

### Model outputs for separated sexes

```{r combined-sex-plot}
x_scale <- scale_x_continuous(
  limits = c(0.25, 1.75),
  breaks = seq(0, 1.75, by = 0.25),
  expand = c(0, 0))

f_weight_plot <- ggplot(
  plotdat_f, 
  aes(
    x = weight_kg, 
    y = fit)) +
  
  geom_line(size = 1) +
  
  geom_ribbon(aes(
    ymin  = lower, 
    ymax  = upper),
    alpha = 0.6,
    fill  = "maroon") +
  
  x_scale +
  
  labs(
    title = "A)", 
    x     = NULL, 
    y     = NULL) +
  
  theme_classic() +
  theme(
    axis.line   = element_line(colour = "black"),
    panel.grid  = element_blank(),
    axis.text   = element_text(color = "black"),
    plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
    plot.title  = element_text(face = "bold", size = 12),
    text        = element_text(size = 12))
f_weight_plot
```

```{r}
plotbreed_m <- ggplot(
  plotdat_m, aes(
    x    = weight_m, 
    y    = fit_response_m,
    ymin = lower_response_m, 
    ymax = upper_response_m)) +
  
  geom_ribbon(
    alpha    = 0.3,
    colour   = "darkgreen",
    fill     = "cadetblue1",
    linetype = "dotted") +
 
  geom_line(colour = "black", linewidth = 1) +
  
  x_scale +
  
  labs(title = "B)", x = "Weight (kilograms)", y = NULL) +
  
  theme_classic() +
  theme(
    axis.line   = element_line(colour = "black"),
    panel.grid  = element_blank(),
    axis.text   = element_text(color = "black"),
    plot.title  = element_text(face = "bold", size = 12),
    plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
    text        = element_text(size = 12))
plotbreed_m
```

```{r}
combined_figure_sexes <- (f_weight_plot / plotbreed_m) +
  plot_annotation(
    title   = "Factors Affecting Breeding Success",
    theme   = theme(plot.title = element_text(
      size  = 16,
      face  = "bold",
      hjust = 0.5)))
combined_figure_sexes

combined_grob <- patchworkGrob(combined_figure_sexes)

y_label <- textGrob(
  "Predicted likelihood of breeding success",
  rot   = 90,
  gp    = gpar(fontsize = 14),
  vjust = 1)

final_figure <- grid.arrange(
  y_label, 
  combined_grob,
  ncol   = 2,
  widths = c(0.05, 0.95))

ggsave(
  "Factors__breeding_success+sexes.png", 
  final_figure,
  width  = 8, 
  height = 6, 
  dpi    = 300)
```

# **Diagnostics**

```{r model-diagnostics}
# Helper: create a formatted table grob with a title
create_table_grob <- function(data, title) {
  table_grob <- tableGrob(
    data,
    rows        = NULL,
    theme       = ttheme_default(
      base_size = 8,
      core      = list(fg_params = list(hjust = 0, x = 0.05),
                       padding   = unit(c(1.5, 1.5), "mm")),
      colhead   = list(fg_params = list(hjust = 0, x = 0.05),
                       padding   = unit(c(1.5, 1.5), "mm"))))

  title_grob <- textGrob(
    title,
    gp   = gpar(fontsize = 9, fontface = "bold"),
    just = "left", 
    x    = 0.02)

  # Prepend title row above the table grob
  combined <- gtable_add_rows(
    table_grob, heights = unit(0.5, "cm"), pos = 0)
  combined <- gtable_add_grob(
    combined, title_grob, t = 1, l = 1, r = ncol(combined))
  return(combined)
}
```

```{r}
# Extract and round collinearity results for each model
binom_collin   <- as.data.frame(check_collinearity(topmod_pruned))
poisson_collin <- as.data.frame(check_collinearity(psn_success))
fem_collin     <- as.data.frame(check_collinearity(binmod_f))
male_collin    <- as.data.frame(check_collinearity(topmod_m))

binom_collin[sapply(binom_collin, is.numeric)] <-
  round(binom_collin[sapply(binom_collin, is.numeric)], 3)

poisson_collin[sapply(poisson_collin, is.numeric)] <-
  round(poisson_collin[sapply(poisson_collin, is.numeric)], 3)

fem_collin[sapply(fem_collin, is.numeric)] <-
  round(fem_collin[sapply(fem_collin, is.numeric)], 3)

male_collin[sapply(male_collin, is.numeric)] <-
  round(male_collin[sapply(male_collin, is.numeric)], 3)
```

```{r}
# Collinearity table grobs
col_table1 <- create_table_grob(
  binom_collin, 
  "Binomial model of breeding success - Collinearity")

col_table2 <- create_table_grob(
  poisson_collin, 
  "Poisson model of individual fecundity - Collinearity")

col_table3 <- create_table_grob(
  fem_collin, 
  "Binomial model of female breeding success - Collinearity")

col_table4 <- create_table_grob(
  male_collin, 
  "Binomial model of male breeding success - Collinearity")
```

```{r}
# Overdispersion table grobs
binom_overdisp1 <- check_overdispersion(topmod_pruned)
overdisp_df1    <- data.frame(
  Metric = c("Dispersion ratio", "p-value"),
  Value  = c(
    round(binom_overdisp1$dispersion_ratio, 3),
    round(binom_overdisp1$p_value, 3)))
od_table1 <- create_table_grob(overdisp_df1, "Overdispersion")

psn_overdisp <- check_overdispersion(psn_success)
overdisp_df2 <- data.frame(
  Metric = c("Dispersion ratio", "Pearson's Chi-Squared", "p-value"),
  Value  = c(
    round(psn_overdisp$dispersion_ratio, 3),
    round(psn_overdisp$chisq, 3),
    round(psn_overdisp$p_value, 3)))
od_table2 <- create_table_grob(overdisp_df2, "Overdispersion")

fem_overdisp <- check_overdispersion(binmod_f)
overdisp_df3 <- data.frame(
  Metric = c("Dispersion ratio", "p-value"),
  Value  = c(
    round(fem_overdisp$dispersion_ratio, 3),
    round(fem_overdisp$p_value, 3)))
od_table3 <- create_table_grob(overdisp_df3, "Overdispersion")

male_overdisp <- check_overdispersion(topmod_m)
overdisp_df4  <- data.frame(
  Metric = c("Dispersion ratio", "p-value"),
  Value  = c(
    round(male_overdisp$dispersion_ratio, 3),
    round(male_overdisp$p_value, 3)))
od_table4 <- create_table_grob(overdisp_df4, "Overdispersion")
```

```{r}
# Combine collinearity and overdispersion tables side-by-side per model,
# then stack all four models vertically
combined_plot <- grid.arrange(
  col_table1, od_table1,
  col_table2, od_table2,
  col_table3, od_table3,
  col_table4, od_table4,
  ncol    = 2, 
  nrow    = 4,
  widths  = c(0.80, 0.20),
  heights = c(1, 1, 1, 1))

ggsave(
  "model_diagnostics_combined.png", 
  combined_plot,
  width  = 210, 
  height = 120, 
  units  = "mm", 
  dpi    = 300, 
  bg     = "white")
```

# **Session information**

```{r, eval = TRUE}
# Display R version, OS, and loaded packages
sessionInfo()
```

```{r, eval = FALSE}
# Generate citation file in /metadata
generate_citation()
```
